\pkg{DNest4}: Diffusive Nested Sampling in \proglang{C++} and \proglang{Python}

### 2016-06-08 paper/ms.tex
RJObject stuff, continued...

To implement this model for \pkg{DNest4}, the model
class needs to contain an instance of an
\code{RJObject}, which contains the parameters
$\{(\mu_j, \sigma_j, w_j)\}$:
\begin{CodeChunk}
\begin{CodeInput}
class MyModel
{
    private:
        DNest4::RJObject<MyConditionalPrior> gaussians;
....
\end{CodeInput}
\end{CodeChunk}
where the template argument \code{<MyConditionalPrior>}
is a class implementing the prior for the hyperparameters
$\alpha$ and the form of the conditional prior $p(x_i | \alpha)$.
The main advantage of the \code{RJObject} class is that
we do not need to implement any proposals for $\{(\mu_j, \sigma_j, W_j)\}$.
Rather, these can be done trivially as follows:
\begin{CodeChunk}
\begin{CodeInput}
    // Generate the components from the prior
    gaussians.from_prior(rng);

@@

The \code{RJObject} constructor definition is:
\begin{CodeChunk}
\begin{CodeInput}
RJObject(int num_dimensions, int max_num_components, bool fixed,
         const ConditionalPrior& conditional_prior,
         PriorType prior_type=PriorType::uniform);
\end{CodeInput}
\end{CodeChunk}
where \code{num\_dimensions} is the number of parameters needed to
specify a single component (three in this example), \code{max\_num\_components}
is the maximum value of $N$ allowed, \code{fixed} determines whether
$N$ is fixed at $N_{\rm max}$ or allowed to vary,
\code{conditional\_prior} is an instance of a conditional prior,
and \code{prior\_type} controls the prior for $N$ (the default is uniform
from $\{0, 1, 2, ..., N_{\rm max}\}$). The {\tt RJObject\_1DMixture}
\code{MyModel} initializes its {\tt RJObject} as follows:
\begin{CodeChunk}
\begin{CodeInput}
MyModel::MyModel()
:gaussians(3, 100, false, MyConditionalPrior(), PriorType::log_uniform)
{
}
\end{CodeInput}
\end{CodeChunk}
Passing \code{PriorType::log_uniform} for the final argument
specifies the $1/(N+1)$ prior for $N$.

The data and the posterior mean fit are shown
in Figure~\ref{fig:galaxies}.

>>>>

\subsection{Using the RJObject class}
To implement this model for \pkg{DNest4}, the model
class needs to contain an instance of an
\code{RJObject}, which contains the parameters
$\{(\mu_j, \sigma_j, w_j)\}$:
\begin{CodeChunk}
\begin{CodeInput}
class MyModel
{
    private:
        DNest4::RJObject<MyConditionalPrior> gaussians;
....
\end{CodeInput}
\end{CodeChunk}
where the template argument \code{<MyConditionalPrior>}
is a class implementing the prior for the hyperparameters
$\alpha$ and the form of the conditional prior $p(x_i | \alpha)$.
The main advantage of the \code{RJObject} class is that
we do not need to implement any proposals for $\{(\mu_j, \sigma_j, W_j)\}$.
Rather, these can be done trivially as follows:
\begin{CodeChunk}
\begin{CodeInput}
    // Generate the components from the prior
    gaussians.from_prior(rng);

@@

The \code{RJObject} constructor definition is:
\begin{CodeChunk}
\begin{CodeInput}
RJObject(int num_dimensions, int max_num_components, bool fixed,
         const ConditionalPrior& conditional_prior,
         PriorType prior_type=PriorType::uniform);
\end{CodeInput}
\end{CodeChunk}
where \code{num\_dimensions} is the number of parameters needed to
specify a single component (three in this example), \code{max\_num\_components}
is the maximum value of $N$ allowed, \code{fixed} determines whether
$N$ is fixed at $N_{\rm max}$ or allowed to vary,
\code{conditional\_prior} is an instance of a conditional prior,
and \code{prior\_type} controls the prior for $N$ (the default is uniform
from $\{0, 1, 2, ..., N_{\rm max}\}$). The {\tt RJObject\_1DMixture}
\code{MyModel} initializes its {\tt RJObject} as follows:
\begin{CodeChunk}
\begin{CodeInput}
MyModel::MyModel()
:gaussians(3, 100, false, MyConditionalPrior(), PriorType::log_uniform)
{
}
\end{CodeInput}
\end{CodeChunk}
Passing \code{PriorType::log_uniform} for the final argument
specifies the $1/(N+1)$ prior for $N$.
One complication for this model is that \code{RJObject}, by default,
allows $N$ to be zero, which makes no sense for this particular
problem. Therefore, we prevent $N=0$ from being generated in
\code{MyModel::from\_prior}, and assert that it should always be
rejected if proposed in \code{MyModel::perturb}:

\begin{CodeChunk}
\begin{CodeInput}
void MyModel::from_prior(RNG& rng)
{
    do
    {
        gaussians.from_prior(rng);
    }while(gaussians.get_components().size() == 0);
}

double MyModel::perturb(RNG& rng)
{
    double logH = 0.0;

    logH += gaussians.perturb(rng);
    if(gaussians.get_components().size() == 0)
        return -std::numeric_limits<double>::max();

    return logH;
}
\end{CodeInput}
\end{CodeChunk}

\subsection{Mixture model results}
The data and the posterior mean fit are shown
in Figure~\ref{fig:galaxies} and the posterior distribution for
$N$ is shown in Figure~\ref{fig:galaxies_N}.
